\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-07-01 tex/paper.tex
save

Complex computer simulators are increasingly used across fields of science to
describe generative models tying parameters of an underlying theory to
experimental observations. Inference in this setup is often
difficult, as simulators rarely provide a way to directly evaluate the likelihood
function for a given observation. In this note, we develop a likelihood-free
inference algorithm for fitting a forward non-differentiable generative model to
continuous or discrete observed data. We adapt the adversarial training procedure of generative
adversarial networks by replacing the implicit generative network with a
domain-based scientific simulator, and solve the resulting non-differentiable
minimax problem by minimizing variational upper bounds of the adversarial objectives.
Effectively, the procedure results in learning an arbitrarily tight
proposal distribution over simulator parameters, such that the corresponding
marginal distribution of the generated data matches the observations.
Preliminary results with non-differentiable simulators producing discrete/continuous univariate/multi-dimensional data
validate the proposed method.

@@

Generative adversarial networks (GANs) were first proposed by
\cite{goodfellow2014generative} as a way to build an implicit generative model
capable of producing samples from random noise $\mathbf{z}$. More specifically,
a generative model $g(\cdot; \theta)$ is pit against an adversarial
classifier $d(\cdot; \phi):\mathbb{R}^d \to [0,1]$ with parameters $\phi$ and whose antagonistic objective is to recognize real data $\mathbf{x}$
from generated data $\tilde{\mathbf{x}} = g(\mathbf{z}; \theta)$. Both models $g$ and $d$
are trained simultaneously, in such a way that $g$ learns to fool
its adversary $d$ (which happens when $g$ produces samples comparable to the
observed data), while $d$ continuously adapts to changes in $g$. When $d$ is
trained to optimality before each parameter update of the generator, it can
be shown that the original adversarial learning procedure amounts to minimizing
the Jensen-Shannon divergence $\text{JSD}(p_r(\mathbf{x}) \parallel p(\mathbf{x}|\theta))$ between $p_r(\mathbf{x})$ and $p(\mathbf{x}|\theta)$.

@@

Under mild restrictions outlined in  \citep{2012arXiv1212.4507S}, the bound
$U(\psi)$ is differentiable with respect to $\psi$, and using the log-likelihood
trick it comes:
\begin{align}\label{eqn:approx-grad}
    \nabla_\psi U(\psi) &= \nabla_\psi \mathbb{E}_{\theta \sim q(\theta|\psi)} [f(\theta)] \nonumber \\
    &= \nabla_\psi \int f(\theta)  q(\theta|\psi)  d\theta \nonumber \\
    &= \int f(\theta) \nabla_\psi q(\theta|\psi)  d\theta \nonumber \\
    &= \int \left[ f(\theta) \nabla_\psi \log q(\theta|\psi) \right]  q(\theta|\psi) d\theta \nonumber \\
    &= \mathbb{E}_{\theta \sim q(\theta|\psi)} [f(\theta) \nabla_\psi \log q(\theta|\psi)]
\end{align}
Effectively, this means that provided that the score function $\nabla_\psi \log
q(\theta|\psi)$ of the proposal is known and that one can evaluate
$f(\mathbf{\theta})$ for any $\theta$, then one can construct empirical
estimates of Eqn.~\ref{eqn:approx-grad}, which can in turn be used to minimize
$U(\psi)$ with stochastic gradient descent (or a variant thereof, like
Adam~\cite{2014arXiv1412.6980K} or the Natural Evolution Strategy
algorithm~\citep{2011arXiv1106.4487W}, for scaling invariance and
robustness to noisy gradients).

@@

\begin{figure*}
    \begin{minipage}{\linewidth}
    \begin{algorithm}[H]
    \caption{Adversarial variational optimization.}

@@

The alternating stochastic gradient descent on ${\cal L}_d$ and ${\cal L}_g$ in
GANs (Section~\ref{sec:gans}) inherently assumes that the generator $g$ is a differentiable function. In
the setting where we are not interested in learning the implicit model itself but are
rather interested in the inference of parameters of a fixed
non-differentiable simulator (Section~\ref{sec:problem}),
gradients $\nabla_\theta g$ either do not exist or cannot be accessed. As a
result, gradients $\nabla_\theta {\cal L}_g$ cannot be constructed and the
optimization procedure cannot be carried out.

In this work, we propose to rely on variational optimization to minimize ${\cal
L}_d$ and ${\cal L}_g$, thereby bypassing the non-differentiability of $g$. More
specifically, we consider a proposal distribution $q(\theta|\psi)$ over the
parameters of $g$ and $p(\mathbf{x}|\theta)$ and minimize in alternation the variational upper bounds
\begin{align}
    U_d &= \mathbb{E}_{\theta \sim q(\theta|\psi)} [ {\cal L}_d ] \label{eqn:vo-ud} \\
    U_g &= \mathbb{E}_{\theta \sim q(\theta|\psi)} [ {\cal L}_g ] \label{eqn:vo-ug}
\end{align} respectively over $\phi$ and $\psi$.
When updating
$\phi$, unbiased estimates of $\nabla_\phi U_d$ can be obtained by
evaluating the exact and known gradient of $U_d$ over mini-batches of true and
generated data, as ordinarily done in stochastic gradient descent. When updating
$\psi$, estimates of $\nabla_\psi U_g$ can be derived with forward
simulations, as described in the previous section.
That is,
\begin{equation}\label{eqn:grad-ug-approx}
    \nabla_\psi U_g = \mathbb{E}_{\theta \sim q(\theta|\psi), \mathbf{z} \sim p(\mathbf{z}|\theta)}  [-d(g(\mathbf{z};\theta);\phi) \nabla_\psi \log q(\theta|\psi)],
\end{equation} which we can approximate with mini-batches of
generated data
\begin{equation}
    \nabla_\psi U_g \approx \frac{1}{M} \sum_{m=1}^M -d(g(\mathbf{z}_m; \theta_m); \phi) \nabla_\psi \log q(\theta_m|\psi)
\end{equation}
for $\theta_m \sim q(\theta|\psi)$ and $\mathbf{z}_m \sim p(\mathbf{z}|\theta_m)$.
For completeness, Algorithm~\ref{alg:avo} outlines the proposed adversarial variational
optimization procedure, as built on top of WGAN-GP.
Obviously, the variational relaxation could similarly be coupled with
other variants of GANs and/or of evolution strategies.

@@

As a first illustrative experiment, we evaluate inference for a discrete Poisson
distribution with unknown scale parameter $\lambda$. We artificially consider
the distribution as a parameterized simulator, from which we can only
generate data.

The observed data is sampled from a Poisson with scale $\lambda^* = 7$.
Algorithm~\ref{alg:avo} is run for 300 epochs with mini-batches of size $M=64$
and the following configuration. For the critic $d$, we use a 3-layer MLP with 10
hidden nodes per layer and ReLU activations. At each epoch, Adam is run for
$n_\text{critic}=100$ iterations with a step size $\alpha=0.01$, decay rates
$\beta_1=\beta_2=0.5$ and its inner first and second moment vectors reset at
each outer epoch in order to avoid building momentum in staled directions.  For
estimating $\lambda^*$, we parameterize $\theta$ as $\log(\lambda)$ and use a univariate Gaussian proposal distribution
$q(\theta|\psi)$ initialized with a mean at $\log(5)$ and unit variance. At
each epoch, parameters $\psi$ are updated by taking one Adam step, with
$\alpha=0.01$ and $\beta_1=\beta_2=0.5$. The gradient penalty coefficient is set to
$\lambda=0.025$, and the entropy penalty is evaluated at both $\gamma=0$ and $\gamma=5$.

\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{figures/poisson.pdf}
    \caption{Discrete Poisson model with unknown scale.
             ({\it Top left}) Proposal distributions $q(\theta|\psi)$ after adversarial variational optimization. For both $\gamma=0$ and $\gamma=5$, the distributions correctly concentrate their density around
                        the true value $\log(\lambda^*)$. Penalizing the entropy of the proposal distribution ($\gamma=5$) results in a tighter density and therefore to better inference.
             ({\it Top right}) Model distributions $p(\mathbf{x}|\psi)$ after training. This plot shows that the resulting parameterizations of the simulator closely reproduce the true distribution.
             ({\it Bottom}) Empirical estimates of the variational upper bound $U_d$ as optimization progresses.
             }\label{fig:poisson}
\end{figure}

The top left plot in Figure~\ref{fig:poisson} illustrates the resulting proposal
distributions $q(\theta|\psi)$ after adversarial variational optimization.  For
both $\gamma=0$ and $\gamma=5$, the proposal distributions correctly concentrate
their density around the true parameter value $\log(\lambda^*) = 1.94$. Under
the effect of the positive entropy penalty $H(q(\theta|\psi))$,
the proposal distribution for $\gamma=5$ concentrates its mass more tightly,
yielding in this case a more precise inference.  The top right plot compares the
model distributions to the true distribution.  As theoretically expected from
adversarial training, we see that the resulting distributions closely match
the true distribution, with in this case slightly  better results for the penalized
model.  The bottom plot of Figure~\ref{fig:poisson} shows empirical estimates
of $-U_d$ with respect to the epoch number. For both $\gamma=0$ and $\gamma=5$,
the curves quickly fall towards $0$, which indicates that
$\mathbb{E}_{\tilde{\mathbf{x}} \sim p(\mathbf{x}|\theta)}
[d(\tilde{\mathbf{x}};\phi)] \approx \mathbb{E}_{\mathbf{x} \sim
p_r(\mathbf{x})} [d(\mathbf{x};\phi)]$ and that the critic cannot distinguish
between true and model data. Despite the discreteness and the
non-differentiability of the underlying generator, this confirms that inference
with adversarial variational optimization works.

@@

As a second example, we evaluate parameter inference for a generator producing
5-dimensional continuous data, as originally specified in Section 4.2. of
\citep{cranmer2015approximating}. More specifically, we consider the following
generative process:
\begin{itemize}
    \item $\mathbf{z} = (z_0, z_1, z_2, z_3, z_4)$, such that
        $z_0 \sim {\cal N}(\mu=\alpha, \sigma=1)$,
        $z_1 \sim {\cal N}(\mu=\beta, \sigma=3)$,
        $z_2 \sim {\text{Mixture}}(\frac{1}{2}\,{\cal N}(\mu=-2, \sigma=1), \frac{1}{2}\,{\cal N}(\mu=2, \sigma=0.5))$,
        $z_3 \sim {\text{Exponential}(\lambda=3)}$, and
        $z_4 \sim {\text{Exponential}(\lambda=0.5)}$;
    \item $\mathbf{x} = R  \mathbf{z}$, where $R$ is a fixed
    semi-positive definite $5 \times 5$ matrix defining a fixed projection
    of $\mathbf{z}$ into the observed space.
\end{itemize}
We consider observed data generated at the nominal values $\theta^* = (\alpha^*=1,\beta^*=-1)$.
The simulator parameters are modeled with a factored Gaussian
proposal distribution $q(\theta|\psi) = q(\alpha|\psi) q(\beta|\psi)$, where each component is
initialized with zero mean and unit variance.
Hyper-parameters are set to $M=64$, $n_\text{critic}=100$, $\lambda=0.025$, $\gamma=10$ and
Adam configured with $\alpha=0.01$, $\beta_1=0.9$ and $\beta_2=0.999$.
The network architecture of the critic is the same as in the previous example.

@@

The top left plot in Figure~\ref{fig:weinberg} illustrates the resulting
proposal distributions $q(\theta|\psi)$ for $\gamma=0$ and $\gamma=5$ after
adversarial variational optimization. We see that the distributions both arrive
in the neighborhood of $\theta^*$, with a density highly more concentrated for
$\gamma=5$ than for $\gamma=0$.  Despite these differences and the relative
distance with $\theta^*$, both models closely match the observed data, as shown
in the top right plot of  Figure~\ref{fig:weinberg}, with again slightly better
results for the entropy penalized model. This suggests either a relatively flat
landscape around $\theta^*$ or that the observed data can in this case also be
reproduced with a smeared variation $p(\mathbf{x}|\psi)$ of the simulator.
Finally, the bottom plot of Figure~\ref{fig:weinberg} shows that for both
$\gamma=0$ and $\gamma=5$ the variational upper bound $-U_d$ quickly fall
towards $0$, which indicates  convergence towards a distribution that the critic
cannot distinguish from the true distribution. In summary, despite the intricate
relation betwen the observations $\mathbf{x}$, the latent variable $\mathbf{z}$
and the parameters $\theta$, this example demonstrates the efficacy
of adversarial variational optimization for parameter inference on a black box
simulator.

@@

In this context, the proposed method can be considered as a direct adaptation of
generative adversarial networks~\citep{goodfellow2014generative} to
non-differentiable simulators. It also constitutes an approximate gradient
descent alternative to bayesian optimization based on density
ratios~\citep{cranmer2015approximating,cranmer2016experiments} or to classifier
ABC~\citep{gutmann2017likelihood}.

>>>>

Complex computer simulators are increasingly used across fields of science as
generative models tying parameters of an underlying theory to
experimental observations. Inference in this setup is often
difficult, as simulators rarely provide a way to directly evaluate the likelihood
function. In this note, we develop a likelihood-free
inference algorithm for fitting a forward, non-differentiable generative model to
continuous or discrete observed data. We adapt the adversarial training procedure of generative
adversarial networks by replacing the differentiable generative network with a
domain-specific simulator and solve the resulting non-differentiable
minimax problem by minimizing variational upper bounds of the adversarial objectives.
Effectively, the procedure results in learning an arbitrarily tight
proposal distribution over simulator parameters, such that the corresponding
marginal distribution of the generated data matches the observations.
Preliminary results with non-differentiable simulators producing discrete/continuous univariate/multi-dimensional data
validate the proposed method.

@@

Generative adversarial networks (GANs) were first proposed by
\cite{goodfellow2014generative} as a way to build an implicit generative model
capable of producing samples from random noise $\mathbf{z}$. More specifically,
a generative model $g(\cdot; \theta)$ is pit against an adversarial
classifier $d(\cdot; \phi):\mathbb{R}^d \to [0,1]$ with parameters $\phi$ and whose antagonistic objective is to recognize real data $\mathbf{x}$
from generated data $\tilde{\mathbf{x}} = g(\mathbf{z}; \theta)$. Both models $g$ and $d$
are trained simultaneously, in such a way that $g$ learns to fool
its adversary $d$ (which happens when $g$ produces samples comparable to the
observed data), while $d$ continuously adapts to changes in $g$. When $d$ is
trained to optimality before each parameter update of the generator, it can
be shown that the original adversarial learning procedure~\cite{goodfellow2014generative} amounts to minimizing
the Jensen-Shannon divergence $\text{JSD}(p_r(\mathbf{x}) \parallel p(\mathbf{x}|\theta))$ between $p_r(\mathbf{x})$ and $p(\mathbf{x}|\theta)$.

@@

Under mild restrictions outlined in  \citep{2012arXiv1212.4507S}, the bound
$U(\psi)$ is differentiable with respect to $\psi$, and using the log-likelihood
trick it comes:
\begin{align}\label{eqn:approx-grad}
    \nabla_\psi U(\psi) &= \nabla_\psi \mathbb{E}_{\theta \sim q(\theta|\psi)} [f(\theta)] \nonumber \\
    %&= \nabla_\psi \int f(\theta)  q(\theta|\psi)  d\theta \nonumber \\
    &= \int f(\theta) \nabla_\psi q(\theta|\psi)  d\theta \nonumber \\
    &= \int \left[ f(\theta) \nabla_\psi \log q(\theta|\psi) \right]  q(\theta|\psi) d\theta \nonumber \\
    &= \mathbb{E}_{\theta \sim q(\theta|\psi)} [f(\theta) \nabla_\psi \log q(\theta|\psi)]
\end{align}
Effectively, this means that provided that the score function $\nabla_\psi \log
q(\theta|\psi)$ of the proposal is known and that one can evaluate
$f(\mathbf{\theta})$ for any $\theta$, then one can construct empirical
estimates of Eqn.~\ref{eqn:approx-grad}, which can in turn be used to minimize
$U(\psi)$ with stochastic gradient descent (or a variant thereof, like
Adam~\cite{2014arXiv1412.6980K} or the Natural Evolution Strategy
algorithm~\citep{2011arXiv1106.4487W}, for scaling invariance and
robustness to noisy gradients).

@@

\begin{figure*}
    \begin{minipage}{\linewidth}
    \begin{algorithm}[H]
    \caption{Adversarial variational optimization (AVO).}

@@

The alternating stochastic gradient descent on ${\cal L}_d$ and ${\cal L}_g$ in
GANs (Section~\ref{sec:gans}) inherently assumes that the generator $g$ is a differentiable function. In
the setting where we are interested in the inference of the parameters of a
fixed non-differentiable simulator (Section~\ref{sec:problem}),
rather than in learning the generative model itself,
gradients $\nabla_\theta g$ either do not exist or cannot be accessed. As a
result, gradients $\nabla_\theta {\cal L}_g$ cannot be constructed and the
optimization procedure cannot be carried out.

In this work, we propose to rely on variational optimization to minimize ${\cal
L}_d$ and ${\cal L}_g$, thereby bypassing the non-differentiability of $g$. More
specifically, we consider a proposal distribution $q(\theta|\psi)$ over the
parameters of $g$ and $p(\mathbf{x}|\theta)$ and minimize in alternation the variational upper bounds
\begin{align}
    U_d &= \mathbb{E}_{\theta \sim q(\theta|\psi)} [ {\cal L}_d ] \label{eqn:vo-ud} \\
    U_g &= \mathbb{E}_{\theta \sim q(\theta|\psi)} [ {\cal L}_g ] \label{eqn:vo-ug}
\end{align} respectively over $\phi$ and $\psi$.
When updating
$\phi$, unbiased estimates of $\nabla_\phi U_d$ can be obtained by
evaluating the exact and known gradient of $U_d$ over mini-batches of true and
generated data, as ordinarily done in stochastic gradient descent. When updating
$\psi$, estimates of $\nabla_\psi U_g$ can be derived with forward
simulations, as described in the previous section.
That is,
\begin{equation}\label{eqn:grad-ug-approx}
    \nabla_\psi U_g = \mathbb{E}_{\theta \sim q(\theta|\psi), \mathbf{z} \sim p(\mathbf{z}|\theta)}  [-d(g(\mathbf{z};\theta);\phi) \nabla_\psi \log q(\theta|\psi)],
\end{equation} which we can approximate with mini-batches of
generated data
\begin{equation}
    \nabla_\psi U_g \approx \frac{1}{M} \sum_{m=1}^M -d(g(\mathbf{z}_m; \theta_m); \phi) \nabla_\psi \log q(\theta_m|\psi)
\end{equation}
for $\theta_m \sim q(\theta|\psi)$ and $\mathbf{z}_m \sim p(\mathbf{z}|\theta_m)$.
For completeness, Algorithm~\ref{alg:avo} outlines the proposed adversarial variational
optimization (AVO) procedure, as built on top of WGAN-GP.
Obviously, the variational relaxation could similarly be coupled with
other variants of GANs and/or of evolution strategies.

@@

As a first illustrative experiment, we evaluate inference for a discrete Poisson
distribution with unknown mean $\lambda$. We artificially consider
the distribution as a parameterized simulator, from which we can only
generate data.

The observed data is sampled from a Poisson with mean $\lambda^* = 7$.
Algorithm~\ref{alg:avo} is run for 300 epochs with mini-batches of size $M=64$
and the following configuration. For the critic $d$, we use a 3-layer MLP with 10
hidden nodes per layer and ReLU activations. At each epoch, Adam is run for
$n_\text{critic}=100$ iterations with a step size $\alpha=0.01$, decay rates
$\beta_1=\beta_2=0.5$ and its inner first and second moment vectors reset at
each outer epoch in order to avoid building momentum in staled directions.  For
estimating $\lambda^*$, we parameterize $\theta$ as $\log(\lambda)$ and use a univariate Gaussian proposal distribution
$q(\theta|\psi)$ initialized with a mean at $\log(5)$ and unit variance. At
each epoch, parameters $\psi$ are updated by taking one Adam step, with
$\alpha=0.01$ and $\beta_1=\beta_2=0.5$. The gradient penalty coefficient is set to
$\lambda=0.025$, and the entropy penalty is evaluated at both $\gamma=0$ and $\gamma=5$.

\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{figures/poisson.pdf}
    \caption{Discrete Poisson model with unknown mean.
             ({\it Top left}) Proposal distributions $q(\theta|\psi)$ after adversarial variational optimization. For both $\gamma=0$ and $\gamma=5$, the distributions correctly concentrate their density around
                        the true value $\log(\lambda^*)$. Penalizing the entropy of the proposal distribution ($\gamma=5$) results in a tighter density.
             ({\it Top right}) Model distributions $p(\mathbf{x}|\psi)$ after training. This plot shows that the resulting parameterizations of the simulator closely reproduce the true distribution.
             ({\it Bottom}) Empirical estimates of the variational upper bound $U_d$ as optimization progresses.
             }\label{fig:poisson}
\end{figure}

The top left plot in Figure~\ref{fig:poisson} illustrates the resulting proposal
distributions $q(\theta|\psi)$ after AVO.  For
both $\gamma=0$ and $\gamma=5$, the proposal distributions correctly concentrate
their density around the true parameter value $\log(\lambda^*) = 1.94$. Under
the effect of the positive entropy penalty $H(q(\theta|\psi))$,
the proposal distribution for $\gamma=5$ concentrates its mass more tightly,
yielding in this case a more precise inference.  The top right plot compares the
model distributions to the true distribution.  As theoretically expected from
adversarial training, we see that the resulting distributions closely match
the true distribution, with in this case visually slightly better results for the penalized
model.  The bottom plot of Figure~\ref{fig:poisson} shows empirical estimates
of $-U_d$ with respect to the epoch number. For both $\gamma=0$ and $\gamma=5$,
the curves quickly fall towards $0$, which indicates that
$\mathbb{E}_{\tilde{\mathbf{x}} \sim p(\mathbf{x}|\theta)}
[d(\tilde{\mathbf{x}};\phi)] \approx \mathbb{E}_{\mathbf{x} \sim
p_r(\mathbf{x})} [d(\mathbf{x};\phi)]$ and that the critic cannot distinguish
between true and model data. Despite the discreteness and the
non-differentiability of the underlying generator, this confirms that inference
with adversarial variational optimization works.

@@

As a second toy example, we evaluate parameter inference for a generator producing
5-dimensional continuous data, as originally specified in Section 4.2. of
\citep{cranmer2015approximating}. More specifically, we consider the following
generative process:
\begin{itemize}
    \item $\mathbf{z} = (z_0, z_1, z_2, z_3, z_4)$, such that
        $z_0 \sim {\cal N}(\mu=\alpha, \sigma=1)$,
        $z_1 \sim {\cal N}(\mu=\beta, \sigma=3)$,
        $z_2 \sim {\text{Mixture}}(\frac{1}{2}\,{\cal N}(\mu=-2, \sigma=1), \frac{1}{2}\,{\cal N}(\mu=2, \sigma=0.5))$,
        $z_3 \sim {\text{Exponential}(\lambda=3)}$, and
        $z_4 \sim {\text{Exponential}(\lambda=0.5)}$;
    \item $\mathbf{x} = R  \mathbf{z}$, where $R$ is a fixed
    semi-positive definite $5 \times 5$ matrix defining a fixed projection
    of $\mathbf{z}$ into the observed space.
\end{itemize}
We consider observed data generated at the nominal values $\theta^* = (\alpha^*=1,\beta^*=-1)$.
The simulator parameters are modeled with a factored Gaussian
proposal distribution $q(\theta|\psi) = q(\alpha|\psi) q(\beta|\psi)$, where each component is
initialized with zero mean and unit variance.
Hyper-parameters are set to $M=64$, $n_\text{critic}=100$, $\lambda=0.025$, $\gamma=10$ and
Adam configured with $\alpha=0.01$, $\beta_1=0.9$ and $\beta_2=0.999$.
The network architecture of the critic is the same as in the previous example.

@@

The top left plot in Figure~\ref{fig:weinberg} illustrates the resulting
proposal distributions $q(\theta|\psi)$ for $\gamma=0$ and $\gamma=5$ after
AVO. We see that the distributions both arrive
in the neighborhood of $\theta^*$, with a density  more highly concentrated for
$\gamma=5$ than for $\gamma=0$.  Despite these differences and the relative
distance with $\theta^*$, both models closely match the observed data, as shown
in the top right plot of  Figure~\ref{fig:weinberg}, with again slightly better
results for the entropy penalized model. This suggests either a relatively flat
landscape around $\theta^*$ or that the observed data can in this case also be
reproduced with a smeared variation $p(\mathbf{x}|\psi)$ of the simulator.
Finally, the bottom plot of Figure~\ref{fig:weinberg} shows that for both
$\gamma=0$ and $\gamma=5$ the variational upper bound $-U_d$ quickly fall
towards $0$, which indicates  convergence towards a distribution that the critic
cannot distinguish from the true distribution. In summary, despite the intricate
relation betwen the observations $\mathbf{x}$, the latent variable $\mathbf{z}$
and the parameters $\theta$, this example demonstrates the efficacy
of AVO for parameter inference on a black box
simulator.

@@

In this context, the proposed method can be considered as a direct adaptation of
generative adversarial networks~\citep{goodfellow2014generative} to
non-differentiable simulators. It also constitutes an approximate gradient
descent alternative to likelihood-free inference based on bayesian optimization and approximated density
ratios~\citep{cranmer2015approximating,cranmer2016experiments} or to classifier
ABC~\citep{gutmann2017likelihood}.
